Online Social Cohesion Reflects Real-World Group Action in Syria during the Arab Spring

### 2020-06-01 manuscript.Rmd
Finish editing Cohesion section

## Social Media and Social Movements
During the Arab Spring, social media provided real-time information about the
events transpiring in countries across the Middle East and North Africa. Our
understanding about the role of social media in social movements is still
evolving. Some argue that these mass movements would not have been possible
without efficient large-scale coordination, both physically on the ground and
virtually on social media platforms.

@@

The words that comprise social media posts are the first level of cohesion that
we will consider. The concept of lexical, or semantic, cohesion has already been
applied to the analysis of political speech (for example, the language of
Margaret Thatcher) and non-political speech [@barzilay1999using;
@morris1991lexical]. In Figure \@ref(fig:cohesion-in-tweets), we see four
sequential tweets, each containing four words.
Each sequential tweet contains some content included in the previous tweets. The
linguistic construct of content word overlap measures the relationship across
units of meaning---here, tweets---where identical and semantically similar words
are repeated [@graesser2004coh]. These repeated lexical
items constitute a theme; on Twitter, this might be a particular hashtag, a
name, a location, or other nominal category. This localized cohesion is
referential; words are repeated and co-referred across tweets, creating a sense
of continuity and cohesiveness of meaning and message.

@@

### Sociolinguistic cohesion: Connectedness of speakers, events, and ideas
We tie the constructs of lexical, interlocutor, and social cohesion
together in a theoretical framework called sociolinguistic cohesion, as shown in
Figure \@ref(fig:theoretical-framework). We can measure the changes in cohesion
preceding and in response to external stimuli by evaluating the messages and the
context in which they occur. Several mechanisms are in motion here. Previous
values of linguistic cohesion should influence future values of linguistic
cohesion. They should also influence future actions in the real world space.
Previous events influence future events, and also the future cohesiveness of the
opposition.

Sociolinguistic cohesion is not a static concept. Just as social movements
experience increased participation as well as setbacks – indicating dynamic
relationships between participants, and participants and the sociopolitical
environment – sociolinguistic cohesion is also dynamic, reflecting the changing
states of the political sphere. Other scholars have noted the diffusion of
messages whereby they reach a critical mass, whereby “behaviors, beliefs, and
preferences” spread within a population [@Barash_Cameron_Macy_2012].  Cohesion
changes over time in response to external events and the internal dynamics of
the social movement. Some may wonder whether social cohesion can diffuse among
distant actors who do not interact face-to-face; Centola and Macy show that
Internet weak ties extend over long distances and as such have spatially
significant abilities to influence actors [@Centola_Macy_2007]. Fink et al. also
explored diffusion of hashtags in viral contagion in Nigeria, a phenomenon
similar to what we identify as sociolinguistic cohesion
[@Fink_Schmidt_Barash_Cameron_Macy_2016]. The prevalence and rate of particular
hashtags contribute to the overall sociolinguistic cohesion. he more that people
reuse hashtags or common phrases, the greater the sociolinguistic cohesion when
this phenomenon occurs during socially significant times like protests
[@Adamic_others_2015]. Exogenous events, like intervention by another country or
international organization like the United Nations, can affect sociolinguistic
cohesion. Endogenous events, like the recruitment of new members to the rebel or
resistance movement, can also affect the level of cohesion.

Combining these separate threads, it is reasonable to expect that real-world
actions and online social interactions would be enmeshed, driving and being
driven by one another. Rather than expecting that only real-world social
cohesion would be driven by real-world action, we would expect interaction
between real-world and virtual dynamics: In this case, exposure to real-world
violence could result in social cohesion in the virtual world. Increased social
cohesion is thus potentially an indicator and a result of increased social
organization and mobilization on the ground.

Exploring how social media influences contentious behavior, Tilly notes that
social media lowers coordination and transaction costs for participants, and
that older versions of organizing social movements have become less effective,
yielding to the new frontier of online organizing [@McAdam_Tarrow_Tilly_2003].
Koopmans extends this further, writing that, “Authorities react to social
movement activities if and as they are depicted in the mass media, and
conversely movement activists become aware of political opportunities and
constraints through the reactions (or non-reactions) that their actions provoke
in the public sphere” [@Koopmans_2004].

Lexical cohesion, then, should be a useful indicator of social cohesion in
political crisis. Lexical cohesion happens at the word level. Interlocutor
cohesion happens between speakers. Political cohesion happens at the social
level during moments of political crisis, like civil wars and social movements.
Sociolinguistic cohesion is comprised of lexical, interlocutor, and
socio-political events, where the words and their context reveal the disposition
of socially connected linguistic communities. Sociolinguistic cohesion combines
the elements of lexical cohesion in that it captures the degree of similarity
between what is said, and how it is said. It accounts for the cohesion between
multiple speakers in an environment whose language evolves to become more
aligned and similar. Sociolinguistic cohesion also addresses how groups involved
in political rebellion must have internal cohesion to overcome the collective
action problem to mount sustained resistance to state authority.

Social coordination has long been used both physically and virtually to start
and carry out collective actions. These collective events bring individuals
together and create emergent patterns of communication and action. Online
communication has historically been used to document and communicate archival
information about events, but it has been increasingly used to bring about
social change. From a dynamical systems perspective, these different means of
social coordination during times of turmoil can be considered two complementary
components of a single system, mutually constrained and constantly interacting
and influencing one another [@mahner1997foundations]. Investigation of the
coupling of real-world mobilization and social cohesion could provide new
insight into the power of social media during mobilization and the ways in which
individuals utilize it to bring about, or respond to, social change.
Increasingly, social media platforms are being utilized during social
mobilization, but the relationship between virtual and real-world mobilization
is still largely unknown. The present work aims to understand how physical and
virtual movements coexist and interact from a dynamical systems perspective.

Just as real-world social cohesion can be quantified, so can online cohesion.
While this can be quantified in many ways, one kind of cohesion can be found in
the similarity in language use on social media. This implicit or explicit
similarity in language use can indicate shared beliefs, thoughts, and feelings,
as has been studied in research on language style in psycholinguistics and
social psychology [@gonzales2010language; @tausczik2010psychological;
@brennan1996conceptual]; although previous work on language style has often
focused on individuals and small groups, this can be scaled up to examine
massive social-level changes [@michel2011quantitative; @fusaroli2015timescales].
This kind of widespread online social cohesion can lead to real-world social
change.

This kind of massive online social cohesion can lead to real-world social
manifestations. Social media platforms provide one way to enable this kind of
online social cohesion by offering distributed information-gathering and
real-time information dissemination. Given its accessibility and proficiency as
an information exchange tool, Twitter has provided individuals involved in
social movements an opportunity to complement or even replace real-world social
mobilization. For example, during the 2012 Italian protests for global economic
change, the riots in Rome were mirrored by Twitter activity both on the day of
the protests and in the days following them [@vicari2013public]. Tweeting was
determined to be more effective locally in discussing real-world, real-time
events than traditional mainstream media platforms. As more and more individuals
rely on social media to obtain local news and information, it becomes
increasingly important to understand just how these micro-blogs reflect
real-world action.

>>>>

## Social Media and Social Movements

Massive online social cohesion can have real-world social
manifestations. Social media platforms provide one way to enable this kind of
online social cohesion by offering distributed information-gathering and
real-time information dissemination. Given its accessibility and proficiency as
an information exchange tool, Twitter has provided individuals involved in
social movements an opportunity to complement or even replace real-world social
mobilization. For example, during the 2012 Italian protests for global economic
change, the riots in Rome were mirrored by Twitter activity both on the day of
the protests and in the days following them [@vicari2013public]. Tweeting was
determined to be more effective locally in discussing real-world, real-time
events than traditional mainstream media platforms. As more and more individuals
rely on social media to obtain local news and information, it becomes
increasingly important to understand just how these micro-blogs reflect
real-world action.

During the Arab Spring, social media provided real-time information about the
events transpiring in countries across the Middle East and North Africa. Our
understanding about the role of social media in social movements is still
evolving. Some argue that these mass movements would not have been possible
without efficient large-scale coordination, both physically on the ground and
virtually on social media platforms.

@@

Just as real-world social cohesion can be quantified, so can online cohesion.
One kind of cohesion can be found in
the similarity in language use on social media. This implicit or explicit
similarity in language use can indicate shared beliefs, thoughts, and feelings,
as has been studied in research on language style in psycholinguistics and
social psychology [@gonzales2010language; @tausczik2010psychological;
@brennan1996conceptual]. Although previous work on language style has often
focused on individuals and small groups, this can be scaled up to examine
massive social-level changes [@michel2011quantitative; @fusaroli2015timescales].

The words that comprise social media posts are the first level of cohesion that
we will consider.  In Figure \@ref(fig:cohesion-in-tweets), we see four
sequential tweets, each containing four words.
Each sequential tweet contains some content included in the previous tweets. The
linguistic construct of content word overlap measures the relationship across
units of meaning---here, tweets---where identical and semantically similar words
are repeated [@graesser2004coh]. These repeated lexical
items constitute a theme; on Twitter, this might be a particular hashtag, a
name, a location, or other nominal category. This localized cohesion is
referential; words are repeated and co-referred across tweets, creating a sense
of continuity and cohesiveness of meaning and message.

@@

### Sociolinguistic Cohesion: Connectedness of Speakers, Ideas, and Events
We tie the constructs of lexical, interlocutor, and social cohesion together in
a theoretical framework called *sociolinguistic cohesion*, as shown in Figure
\@ref(fig:theoretical-framework). Several interconnected mechanisms are in
motion here. Previous values of linguistic cohesion should influence future
values of linguistic cohesion and future actions in the real world space.
Previous events influence future events and future linguistic cohesion in
social media.

Sociolinguistic cohesion is not a static concept. Just as social movements
experience increased participation and setbacks---indicating dynamic
relationships among participants and between participants and the sociopolitical
environment---sociolinguistic cohesion is also dynamic, reflecting the changing
states of the political sphere. Other scholars have noted the diffusion of
messages whereby they reach a critical mass, whereby “behaviors, beliefs, and
preferences” spread within a population [@barash2012critical].  Cohesion
changes over time in response to external events and the internal dynamics of
the social movement. This can occur even if the social interaction occurs entirely 
online [@centola2007complex].

Related previous work has also explored diffusion of hashtags in viral contagion in
Nigeria, a phenomenon similar to what we identify as sociolinguistic cohesion
[@fink2016complex]. The prevalence and rate of particular hashtags were
contributed to the overall sociolinguistic cohesion. The more that people
reuse hashtags or common phrases, the greater the sociolinguistic cohesion when
this phenomenon occurs during socially significant times like protests
[@state2015diffusion].

Combining these separate threads, it is reasonable to expect that real-world
actions and online social interactions would be enmeshed, driving and being
driven by one another. Rather than expecting that only real-world social
cohesion would be driven by real-world action, we would expect interaction
between real-world and online dynamics: In this case, exposure to real-world
violence could result in social cohesion in the virtual world. Increased social
cohesion is thus potentially an indicator and a result of increased social
organization and mobilization on the ground. However, this occurs within a context
that is affected by local and international events; both
exogenous events (like intervention by another country or
international organization like the United Nations) and 
endogenous events (like the recruitment of new members to the rebel or
resistance movement) can affect cohesion.

From a dynamical systems perspective, online social cohesion and
in-person action can be considered two complementary
components of a single system, mutually constrained and constantly interacting
and influencing one another [@mahner1997foundations]. Investigation of the
coupling of real-world mobilization and social cohesion could provide new
insight into the power of social media during mobilization and the ways in which
individuals utilize it to bring about, or respond to, social change.
Increasingly, social media platforms are being utilized during social
mobilization [@mcadam2003dynamics; @koopmans2004movements], 
but the relationship between virtual and real-world mobilization
is still largely unknown. The present work aims to understand how physical and
virtual movements coexist and interact from a dynamical systems perspective.

Lexical cohesion, then, is be a useful indicator of social cohesion in
political crisis. Lexical cohesion happens at the word level. Interlocutor
cohesion happens between speakers. Political cohesion happens at the social
level during moments of political crisis, like civil wars and social movements.
Sociolinguistic cohesion is comprised of lexical, interlocutor, and
socio-political events, where the words and their context reveal the disposition
of socially connected linguistic communities. Sociolinguistic cohesion combines
the elements of lexical cohesion in that it captures the degree of similarity
between what is said, and how it is said. It accounts for the cohesion between
multiple speakers in an environment whose language evolves to become more
aligned and similar. Sociolinguistic cohesion also addresses how groups involved
in political rebellion must have internal cohesion to overcome the collective
action problem to mount sustained resistance to state authority.
